Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-01 manuscript.Rmd
Edits to Lexical and Interlocutor Cohesion sections

The cohesiveness and strength of social movements in the pre-Internet era could be
operationalized by the number of people turning out to protest and through the
political events that precipitated and resulted from social gatherings *en masse*.
The content of placards and protest
signs provided the protest narrative in their own words. Cohesive
social movements were reflected in their cohesive placards and posters. For
example, the ``I am a man'' posters held by the men involved in the 1968
sanitation workers' strike reflected their internal cohesion and unified,
goal-oriented movement [@Windsor2020].

However, the rise of the Internet and its decentralized online networks
has upended theories about recruitment to---and participation in---social
movements. Strong interpersonal ties have yielded to weak impersonal ties that
strongly predict involvement in social movements [@steinert2017spontaneous]. Research
methodologies for measuring social movements have similarly evolved from
ethnographic interviews to ``big data'' analyses of language on social media.

@@

We contextualize linguistic cohesion with social mobilization and response to the relative
intensity of the conflict. Cohesion is a multi-layered concept referring to the
connectedness of words, thoughts, speakers, and actions. Cohesion reflects the
continuity of language and expression, and it shows the connectivity of
participants in a social movement. To address the relationship between language
and discourse and social movements, we adopt an interdisciplinary approach that
incorporates linguistics, psychology, and political science using nonlinear
dynamical systems methodology to model the relationship between language and
action.

### Lexical Cohesion: Connectedness of Words

@@

## Interlocutor cohesion
Lexical cohesion contributes to the next level of relatedness among tweets:
interlocutor cohesion. This is the degree to which people in a conversation,
whether in person or online, have aligned their language [@Coupland_Giles_1988].
Known as communication accommodation theory, the degree of alignment between
speakers reveals their closeness, status, and deliberative strategy
[@Giles_2016]. Speakers can converge and diverge in their communication.
Convergence facilitates easier transmission of information, because the speakers
– or interlocutors – presume some existing familiarity or knowledge base between
them. When you know someone well enough to finish their sentence, this is an
example of convergence. Divergence, on the other hand, is more laborious because
the speakers must make their presumptions explicit and tacit. Interlocutor
cohesion, or convergence, happens when speakers are “singing in the same key” or
are “on the same page.” Windsor (2020) describes various types of cohesion,
using the example of the Occupy Wall Street movement as a symbol of divergence
given the myriad signs, slogans, and interests of the participants
[@Windsor_2020]. On the other hand, the “I am a man” protest captured by Ernest
Withers during the 1968 santiation workers’ strike in Memphis, TN, or the “Black
Lives Matter” placards following the uprising in Ferguson, MO, symbolize
convergence and strong interlocutor cohesion [@Edgar_Johnson_2018].

Cognitive and social psychology theories of alignment and coordination lend
themselves nicely to the study of social media communication during times of
conflict and contentious behavior [@Tollefsen_Dale_Paxton_2013;
@Garrod_Pickering_2004; @Pickering_Garrod_2004]. Over the past few decades, this
area has garnered significant attention and produced evidence showing that
people sometimes linguistically and behaviorally align as they interact more
[@Giles_Coupland_Coupland_1991; @Tollefsen_Dale_Paxton_2013]. Similarly,
accommodation theory refers to the ways in which individuals and groups
linguistically adapt to minimize the social space distance. Speakers in a
collective setting pursue strategies which increase their common ground, and
this is reflected in the language they use to communicate with one another.

A similar process happens among participants in social media discussions. The
process of speakers’ (or tweeters) language becoming more similar is called
linguistic style matching (LSM) [@Niederhoffer_Pennebaker_2002]. LSM is a form
of synchrony between speakers, akin to linguistic convergence. As Bernieri and
Rosenthal (1991) describe, people engaged in linguistic synchrony match each
others’ verbal behaviors, adopting not only similar terms, but similar patterns
of language [@Bernieri_Rosenthal_1991]. For example, LSM is particularly evident
when one speaker introduces swearing into a genteel conversation; the first
usage opens the vulgarity floodgates, and subsequent speakers may repeat the
initial curse, as well as innovate new curses into the conversation. Fed by
lexical cohesion, interlocutor cohesion ebbs and flows along a timeline
contextualized by variation in real world events. Interlocutor cohesion may
increase preceding an in-person social mobilization event, alerting a repressive
government that collective mobilization is happening. Interlocutor cohesion may
also respond to exogenous events such as spontaneous, unprovoked militarized
conflict. If government forces target a new location using military weapons,
this can precipitate online mobilization and discussion of the event.

>>>>

The cohesiveness and strength of social movements in the pre-Internet era could be
operationalized by the number of people turning out to protest and through the
political events that precipitated and resulted from social gatherings *en masse*.
The content of placards and protest
signs provided the protest narrative in their own words. Cohesive
social movements were reflected in their cohesive placards and posters. For
example, the “I am a man” posters held by the men involved in the 1968
sanitation workers' strike reflected their internal cohesion and unified,
goal-oriented movement [@Windsor2020].

However, the rise of the Internet and its decentralized online networks
has upended theories about recruitment to---and participation in---social
movements. Strong interpersonal ties have yielded to weak impersonal ties that
strongly predict involvement in social movements [@steinert2017spontaneous]. Research
methodologies for measuring social movements have similarly evolved from
ethnographic interviews to “big data” analyses of language on social media.

@@

We contextualize linguistic cohesion with social mobilization and response to the relative
intensity of the conflict. Cohesion is a multi-layered concept referring to the
connectedness of speakers, actions, and societies. Cohesion reflects the
continuity of language and expression, and it shows the connectivity of
participants in a social movement. To address the relationship between language
and discourse and social movements, we adopt an interdisciplinary approach that
incorporates linguistics, psychology, and political science using nonlinear
dynamical systems methodology to model the relationship between language and
action.

### Lexical Cohesion: Connectedness of Words between Interlocutors

@@

The use of similar language can provide insight into *interlocutor cohesion*, or
the degree to which people are talking about similar things. This can---as noted
earlier---include people who are talking about similar things from radically
perspectives, as well as those who are talking about something from the same
perspective. Previous research from political science has investigated these
ideas [@Windsor2020]: The “Occupy Wall Street” movement has been held up as an
example of a movement with low interlocutor and linguistic cohesion because of
its wide range of signs and slogans, while the “I am a man” protest (captured by
Ernest Withers during the 1968 sanitation workers' strike in Memphis, TN) or the
“Black Lives Matter” placards (following the uprising in Ferguson, MO) have high
lexical and interlocutor cohesion [@edgar2018struggle].

Fed by lexical cohesion, interlocutor cohesion ebbs and flows along a timeline
contextualized by variation in real world events. Interlocutor cohesion may
increase preceding an in-person social mobilization event, alerting a repressive
government that collective mobilization is happening. Interlocutor cohesion may
also respond to exogenous events such as spontaneous, unprovoked militarized
conflict. If government forces target a new location using military weapons,
this can precipitate online mobilization and discussion of the event.
